---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# epocakir

*/ˈiːpɒk ə keɪ aɪ ɑː/*

<!-- badges: start -->

[![R build status](https://github.com/alwinw/epocakir/workflows/R-CMD-check/badge.svg)](https://github.com/alwinw/epocakir/actions)
[![Codecov test coverage](https://codecov.io/gh/alwinw/epocakir/branch/master/graph/badge.svg)](https://codecov.io/gh/alwinw/epocakir?branch=master)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)

<!-- badges: end -->

## Clinical Coding of Patients with Kidney Disease

The *epocakir* package makes clinical coding of patients with kidney disease using clinical practice guidelines easy.
The guidelines used are the evidence-based [KDIGO guidelines](https://kdigo.org/guidelines/).
This package covers acute kidney injury (AKI), anemia, and chronic liver disease(CKD):

- `aki_staging()`: Classification of AKI staging (`aki_stages`) with automatic selection of:

  - `aki_bCr()`: AKI based on baseline creatinine
  - `aki_SCr()`: AKI based on changes in serum creatinine
  - `aki_UO()`: AKI based on urine output

- `anemia()`: Classification of anemia

- Classification of albuminuria (`Albuminuria_stages`)

  - `Albuminuria_staging_ACR()`: Albuminuria based on Albumin excretion rate
  - `Albuminuria_staging_AER()`: Albuminuria based on Albumin-to-creatinine ratio

- `eGFR()`: Estimation of glomerular filtration rate with automatic selection of:

  - `eGFR_adult_SCr()`: eGFR based on the 2009 CKD-EPI creatinine equation
  - `eGFR_adult_SCysC()`: eGFR based on the 2012 CKD-EPI cystatin C equation
  - `eGFR_adult_SCr_SCysC()`: eGFR based on the 2012 CKD-EPI creatinine-cystatin C equation
  - `eGFR_child_SCr()`: eGFR based on the pediatric creatinine-based equation
  - `eGFR_child_SCr_BUN()`: eGFR based on the pediatric creatinine-BUN equation
  - `eGFR_child_SCysC()`: eGFR based on the pediatric cystatin C-based equation

- `GFR_staging()`: Staging of GFR (`GFR_stages`)
  
- Multiple utility functions including:

  - `conversion_factors`: Conversion factors used throughout the KDIGO guidelines
  - `as_metric()`: Conversion of a measured value into metric units
  - `dob2age()`: Calculation of age from a date of birth
  - `binary2factor()`: Conversion of binary data into factors based on a column name
  - `combine_date_time_cols()`: Combining separate date and time columns into a single date and time column
  - `combn_changes`: Generating changes between measurements


## Installation

You can install the **released** version from [CRAN](https://cran.r-project.org/package=epocakir) with:

```{r cran, eval=FALSE}
install.packages("epocakir")
```

You can install the **development** version from [GitHub](https://github.com/alwinw/epocakir) with:

```{r dev, eval=FALSE}
# install.packages("devtools")
devtools::install_github("alwinw/epocakir")
```

## Getting Started

```{r pkgs, message = FALSE, warning=FALSE}
library(epocakir)
library(dplyr)
library(units)
```

Often clinical data must be cleansed and tidied before analysis can begin.
To assist in this, several utility functions have been included.
To explore these, consider a sample clinical dataset `clinical_obvs`:

```{r clinical_data}
glimpse(clinical_obvs)

tidy_obvs <- clinical_obvs %>%
  combine_date_time_cols() %>%
  mutate(
    Age = dob2age(`Date of Birth`),
    Height = as_metric(height = set_units(as.numeric(Height), "cm"))
  ) %>%
  binary2factor(Male, Surgery)

glimpse(tidy_obvs)
```

Make sure to use `set_units()` from the `units` package to convert all measurements into unit objects for automatic unit conversion in epocakir.

## Examples

It is possible to use `aki_staging()` to automatically classify the presence and staging of AKI.
If a particular method is required, it is possible to classify AKI using `aki_bCr()`, `aki_SCr()` or `aki_UO().`

```{r aki}
head(aki_pt_data)

aki_staging(aki_pt_data,
  SCr = "SCr_", bCr = "bCr_", UO = "UO_",
  dttm = "dttm_", pt_id = "pt_id_"
)

aki_pt_data %>%
  mutate(aki = aki_staging(
    SCr = SCr_, bCr = bCr_, UO = UO_,
    dttm = dttm_, pt_id = pt_id_
  )) %>%
  select(pt_id_, SCr_:dttm_, aki)

aki_pt_data %>%
  mutate(aki = aki_SCr(
    SCr = SCr_, dttm = dttm_, pt_id = pt_id_
  )) %>%
  select(pt_id_, SCr_:dttm_, aki)
```


Similarly, `eGFR()` offers the ability to automatically select the appropriate formula to estimate the glomerular filtration rate.
If a particular formula is required, then `eGFR_adult_SCr`, `eGFR_adult_SCysC`, `eGFR_adult_SCr_SCysC`, `eGFR_child_SCr`, `eGFR_child_SCr_BUN`, or `eGFR_child_SCysC` can be used.

```{r eGFR}
head(eGFR_pt_data)

eGFR(eGFR_pt_data,
  SCr = "SCr_", SCysC = "SCysC_",
  Age = "Age_", height = "height_", BUN = "BUN_",
  male = "male_", black = "black_", pediatric = "pediatric_"
)

eGFR_pt_data %>%
  dplyr::mutate(eGFR = eGFR(
    SCr = SCr_, SCysC = SCysC_,
    Age = Age_, height = height_, BUN = BUN_,
    male = male_, black = black_, pediatric = pediatric_
  )) %>%
  select(SCr_:pediatric_, eGFR)

eGFR_pt_data %>%
  dplyr::mutate(eGFR = eGFR_adult_SCr(
    SCr = SCr_, Age = Age_, male = male_, black = black_
  )) %>%
  select(SCr_:pediatric_, eGFR)
```

## References

KDIGO Guidelines - <https://kdigo.org/guidelines/>

## Activity

![GitHub commit activity](https://img.shields.io/github/commit-activity/m/alwinw/epocakir)
![GitHub last commit](https://img.shields.io/github/last-commit/alwinw/epocakir)
![Visits](https://badges.pufler.dev/visits/alwinw/epocakir?&label=visits)
![GitHub repo size in bytes](https://img.shields.io/github/repo-size/alwinw/epocakir)
![Total Lines](https://img.shields.io/tokei/lines/github/alwinw/epocakir)

----

See <https://alwinw.github.io/epocakir/reference/index.html> for more usage details and package reference.
