---
title: "Introduction to epocakir"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to epocakir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 4,
  fig.width = 7
)
```

## Introduction

The epocakir package makes clinical coding of patients with kidney disease using clinical practice guidelines easy. The guidelines used are the evidence-based [KDIGO guidelines](https://kdigo.org/guidelines/). This package covers acute kidney injury (AKI), anemia, and chronic liver disease(CKD).

## Features

-   Classification of AKI staging:

    -   AKI based on baseline creatinine
    -   AKI based on changes in serum creatinine
    -   AKI based on urine output

-   Classification of anemia

-   Classification of albuminuria

    -   Albuminuria based on Albumin excretion rate
    -   Albuminuria based on Albumin-to-creatinine ratio

-   Estimation of glomerular filtration rate:

    -   eGFR based on the 2009 CKD-EPI creatinine equation
    -   eGFR based on the 2012 CKD-EPI cystatin C equation
    -   eGFR based on the 2012 CKD-EPI creatinine-cystatin C equation
    -   eGFR based on the pediatric creatinine-based equation
    -   eGFR based on the pediatric creatinine-BUN equation
    -   eGFR based on the pediatric cystatin C-based equation

-   Multiple utility functions including:

    -   Conversion factors used throughout the KDIGO guidelines
    -   Conversion of a measured value into metric units
    -   Calculation of age from a date of birth
    -   Conversion of binary data into factors based on a column name
    -   Combining seperate date and time columns into a single date and time column
    -   Generating changes between measurements

-   Automatic conversion of units class objects

-   Tidy output allowing seamless integratation with functions from the tidyverse

-   Tidyeval via programming with dplyr

-   Comprehensive tests and coverage

## Example

```{r setup, message = FALSE, warning=FALSE}
library(epocakir)
library(dplyr)
library(units)
```

Often clinical data must be cleansed and tidied before analysis can begin. To assist in this, several utility functions have been included. To explore these, consider a sample clinical dataset `clinical_obvs`

```{r clinical_data}
# Example workflow: clinical_obvs <- read.csv("cohort.csv")
glimpse(clinical_obvs)

tidy_obvs <- clinical_obvs %>%
  combine_date_time_cols() %>%
  mutate(
    Age = dob2age(`Date of Birth`),
    Height = as_metric(height = set_units(as.numeric(Height), "cm"))
  ) %>%
  binary2factor(Male, Surgery)

glimpse(tidy_obvs)
```
